Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-07 surfaces-write-up/surface-write-up.tex
undo Peter's introduction of a strange typographical error
edits to conclusions

\section{Conclusion}
\label{sec:conc}
This paper extends the work of O'Dorney \cite{dorney:canonical} on $\bp^1$ to
$\bp^m$ and Hirzebruch Surfaces, bounding the degrees of generation and
relations of canonical rings of arbitrary divisors $D$.  The methods for
deducing bounds on generators and relations of $\hirz_k$ applied similar
methods to that of O'Dorney's for $\bp^1$.

We can view $\hirz_0$ as $\bp^1\times \bp^1$ and more generally $\hirz_k$ as a
$\bp^1$ bundle over $\bp^1$.  One might hope that similar ideas could be
generalized to all bundles.  Specifically:
\begin{question}
\label{qn:general-product-bundle}
Suppose $X$ and $Y$ are schemes.  Let $D$ be a divisor on $X\times Y$ or more
generally on an $X$ bundles over $Y$.  Can we describe $D$ in terms of divisors
$D_1, \ldots D_n$ on $X$ and $D_1', \ldots d_n'$ on $Y$.  Furthermore, can we
find generators for $R_D$ and its ideal of relations using those of the
$R_{D_i}$'s and $R_{D_j'}$'s?
\end{question}

An immediate obstruction to this is the fact that we critically used the fact
that $\hirz_k$ has well-defined bi-degrees.  Unfortunately, in general $\Pic(X
\times Y) \not \cong \Pic(X) \times \Pic(Y)$.

Alternatively, we can view $\hirz_1$ as $\bp^2$ blown up at a point.  This
suggests the following question:
\begin{question}
\label{qn:general-blowup}
Let $X$ be a scheme.  Let $X'$ be the blowup of $X$ at a point $P$ with a
divisor $D$.  First, is there a way to express $D$ in terms of divisors $D_1,
\ldots, D_n$ on $X$?  Furthermore, is there a way to describe generators and
relations of $R_D$ in terms of those of $R_{D_i}$ for all $i$.
\end{question}

Fortunately, the Picard group of $X$ blown up at a point is the same as that of
$X$ times an additional copy of $\bz$.  Perhaps analyzing the argument from
Lemma ~\ref{thm:hirz-generators-relations} in the case of $\hirz_1$ from the
perspective of adding a copy of $\bz$ to the Picard group, as induced by
the blowup, might suggest a general method for blowups.

Answering Question ~\ref{qn:general-blowup} would in fact provide an answer for
all rational surfaces, via the minimal model program for surfaces.  One might then hope to further extend these results to all minimal surfaces:
\begin{question}
\label{qn:general-minimal-surface}
Can we describe generators and relations of $R_D$ for $D$ a divisor on an
arbitrary minimal surface $X$?
\end{question}

>>>>

\section{Further Questions}
\label{sec:conc}
In this section, we briefly state some further questions.

Recall that every minimal rational surface is either isomorphic to
$\bp^2$ or $\hirz_m$ for some $m \geq 0, m \neq 1$. By Theorems
~\ref{thm:proj-generators-relations} and ~\ref{thm:hirz-generators-relations}, we have given bounds for the generators and relations
of arbitrary canonical rings on any minimal rational surface. A
natural extension of our results is the following.
\begin{question}
\label{qn:general-minimal-surface}
Can we describe generators and relations of $R_D$ for $D$ a divisor on an
arbitrary rational surface $X$?
\end{question}

By the classification of rational surfaces, every rational surface
can be obtained from a minimal rational surface by a sequence of blow-ups \todo{Aaron: Include reference}. Therefore, to answer Question \ref{qn:general-minimal-surface} affirmatively, it suffices
to bound the degree of generators and relations of the 
canonical ring of a divisor on a blow-up of that surface in terms of the canonical ring of a divisor on that surface.

Another direction to generalize the work in this paper
would be to try to express canonical rings of $\bq$-divisors on
$X \times Y$ in terms of $X$ and $Y$. In this paper,
we bounded the degrees of presentations on canonical rings on $\bp^1 \times \bp^1 \cong \hirz_0$.
Perhaps similar techniques can be used to bound
degrees of presentaitons on canonical rings on $(\bp^1)^k$ or
more generally on $(\bp^1)^{i_1} \times \cdots (\bp^k)^{i_k}.$
One might further try to generalize this to bounding degrees of presentations on bundles over $\bp^m$ or on more general
products of schemes.
